Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme\\ \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

### 2020-04-27 LME-MPM.tex
Nuevo pdf
Cambios comentarios PNA

\newcommand{\red}[1]{
  \textcolor{red}{{#1}}
}

@@

@@

\title{Enhanced Material Point Method to face
  dynamic problems: Local-maximum entropy approximation and explicit
  predictor-corrector scheme\\
  \red{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}}

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an Explicit Predictor-Corrector time
  integration scheme has been proposed. Its \textcolor{red}{powerful}
  \cancel{superior} performance mitigates the presence of spurious
  oscillations with minimal dissipation in high frequency
  problems. Other source of numerical noise in \acrshort{mpm} occurs
  when the material points cross computational grid boundaries, being
  this issue motivated by the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the
  material points, where strain-stress fields are computed. This could
  lead to inaccurate solutions as well as aborted simulations in the
  worst cases. To overcome it, this document adopts the \acrfull{lme}
  approximation schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded
  as a \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies. \red{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\cancel{However, this method suffers other kind of instabilities, such those} \red{On the one hand, the main source of instability occurs}
when material points cross cell boundaries. This \cancel{give rise to} provoked the
development of other interpolation techniques to overcome this
limitation such as the \acrfull{gimp} method Bardenhagen \& Kober
(2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}. Therefore \acrshort{ddmp} a large number of particles is needed for convergence, what
makes the method very expensive. In recent years the employment of
spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread in
a challenging way, the quality of the \acrshort{ctls} approximation decreases locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute \cancel{for the
wide range of} \red{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas {\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and coworkers~\cite{Wriggers18} with implicit schemes. And more
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}. \cancel{but without exploring} \red{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing \red{somehow} the accuracy. \cancel{of
  the time integration scheme.} \MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach, and introducing the well-know
\acrfull{igimp} method. However, the local damping introduced by
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.}
In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several \cancel{simple but} challenging test cases \red{on the elastic regime}: the
propagation of shock waves in a elastic bar and the response of a block of soil gradually loaded with gravitational forces.

@@

\red{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. \red{First (i) the \textit{compatibility
  equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type \gls{dirichlet-boundary}. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. \red{Therefore, the stress field \gls{stress} will be considered the corresponding conjugate variable for the strain field, being the one} which satisfies (ii) the \textit{conservation of
  momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor:
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\red{In this research, plane strain Linear Elasticity has been considered. Thus the constitutive tensor, \tens{D}, is the well known linear elastic one.} The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

@@

In order to obtain a finite set of equations, in contrast with the
\acrshort{fem}, in \acrshort{mpm} a double discretization procedure is
performed. First, the continuum \gls{domain} is discretized with a finite sum of material points (also denominated particles the manuscript). Each material point
represents a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point location, $\vec{x}_p$, is defined at the centroid
of each $\Omega_p$ (see figure~\ref{fig:MPM-discretization} for details).
Initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point, which also owns the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where Riemann \red{ALGUNA CITA AQUI}
integral definition is recovered as an addition of a finite set of points, and
their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Following, the aforementioned second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated, being $N_n$ the number of grid
nodes. \red{Spatial derivatives, such as gradients and
divergences, are computed through the support of the background mesh.}

>>>>

\newcommand{\PNA}[1]{
  \textcolor{red}{{#1}}
}

@@

\newcommand{\MMS}[1]{
  \textcolor{gree}{{#1}}
}

@@

\title{Enhanced Material Point Method to face
  dynamic problems: Local-maximum entropy approximation and explicit
  predictor-corrector scheme\\
  \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}}

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an Explicit Predictor-Corrector time
  integration scheme has been proposed. Its \textcolor{red}{powerful}
  performance mitigates the presence of spurious oscillations with
  minimal dissipation in high frequency problems. Other source of
  numerical noise in \acrshort{mpm} occurs when the material points
  cross computational grid boundaries, being this issue motivated by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  overcome it, this document adopts the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies. \PNA{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\PNA{On the one hand, the main source of instability occurs} when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}. Therefore \acrshort{ddmp} a large number of particles is needed for convergence, what
makes the method very expensive. In recent years the employment of
spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread in
a challenging way, the quality of the \acrshort{ctls} approximation decreases locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute \PNA{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas {\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and coworkers~\cite{Wriggers18} with implicit schemes. And more
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}.\PNA{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing \PNA{somehow} the accuracy. \MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several challenging test cases \PNA{on the elastic regime}: the
propagation of shock waves in a elastic bar and the response of a block of soil gradually loaded with gravitational forces.

@@

\PNA{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. \PNA{First (i) the \textit{compatibility
  equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type \gls{dirichlet-boundary}. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. \PNA{Therefore, the stress field \gls{stress} will be considered the corresponding conjugate variable for the strain field, being the one} which satisfies (ii) the \textit{conservation of
  momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor:
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\PNA{In this research, plane strain Linear Elasticity has been considered. Thus the constitutive tensor, \tens{D}, is the well known linear elastic one.} The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

@@

In order to obtain a finite set of equations, in contrast with the
\acrshort{fem}, in \acrshort{mpm} a double discretization procedure is
performed. First, the continuum \gls{domain} is discretized with a finite sum of material points (also denominated particles the manuscript). Each material point
represents a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point location, $\vec{x}_p$, is defined at the centroid
of each $\Omega_p$ (see figure~\ref{fig:MPM-discretization} for details).
Initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point, which also owns the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where Riemann \PNA{ALGUNA CITA AQUI}
integral definition is recovered as an addition of a finite set of points, and
their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Following, the aforementioned second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated, being $N_n$ the number of grid
nodes. \PNA{Spatial derivatives, such as gradients and
divergences, are computed through the support of the background mesh.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Mesh-particles-back}}
  \caption{Description of the spatial discretization for domain presented in the
    figure~\ref{fig:Continuum-solid}. Blue mesh represent the
    background computational support, and the red mesh conforms the
    discretized continuum body.}
  \label{fig:MPM-discretization}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces},  \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
approximating the displacement field of the particle
$p$ as $\vec{u}_{p} = N_{Ip} \vect{u}_I$, $\vec{u}^{\psi}_{p} = N_{Ip} \vect{u}_I^{\psi}$,
and its gradient as $\tens{\varepsilon}_p = (\vect{u}_I \otimes
\Grad{N_{Ip}})^s$, $\tens{\varepsilon}_p^{\psi}= (\vect{u}_I^{\psi} \otimes \Grad{N_{Ip}})^s$.
nodal balance of forces of the continuum yields:
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$ and $\tens{m}_{IJ}$, the nodal mass matrix, is obtained through:
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} = N_{Ip} m_p N_{Jp}.
\end{equation}
In order to improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_{IJ}^{lumped}$.
Following, internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \tens{\sigma}_{p} \cdot \Grad{N_{Ip}} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = N_{Ip}\ \vec{b}_{p}\ m_p  + N_{Ip}\ \vec{t}^s_{p}\ m_p h^{-1} 
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor rate, $\dot{ \tens{\varepsilon}}_{p}$, as a measure of the time derivative of the strain tensor, is updated employing the velocity at the background mesh by the equation:
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2} \left[\Grad{N_{Ip}}\ \otimes \vec{v}_{I} + \vec{v}_{I} \otimes
    \Grad{N_{Ip}}\ \right].
\end{equation}
Next, mass conservation is guaranteed by enforcing the null value of
the material derivative of the density field $\frac{D \rho}{D t} = 0$.
This leads to a suitable equation to update the density field:
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{trace} \left( \dot{\tens{\varepsilon}} \right).
\end{equation}
Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order \PNA{time} integration scheme is required. Therefore, time is
discretized into a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation is solved, the values at the nodes are
interpolated back into the particles, which are advected
to the new position through:
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = N_{Ip}\ \vec{v}_{I}  
\end{equation}
\PNA{Traditionally, Eqs.
\eqref{eq:particle_balance_forces3} and ~\eqref{eq:Updated_Lagrangian},
are solved with an explicit forward Euler algorithm. In the following subsection, this and the proposed schemes are described}.
